Stack Safety is a Security Property

### 2021-04-08 tex/main.tex
Further progress on examples

\section{Stack Safety, Intuitively}
\label{sec:running-example}
\colorlet{lgray}{gray!40}

\newcommand{\mainsealc}{cyan}
\newcommand{\fsealc}{green}
\newcommand{\unsealc}{lgray}
\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
\newcommand{\fulloutc}{white}

@@

\Cref{fig:int1} shows an integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}, performs arithmetic, and returns a result.
In the initial memory layout the only frame is currently {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

\sna{I'm experimenting with the example style, and it ends here currently, so below
  this is the old version in which the stack is contiguous. Current plan is to switch
  to contiguous stack when adding stack passing.}

@@

\begin{subfigure}{.33\textwidth}
\begin{verbatim}
int main() {
0:     int x = 42, y = 0;
1[C]:  y = f();
2:     print (x + y);
}

int f() {
10:     int z = *($SP-1);
11[R]:  return (z+z);
}
\end{verbatim}
\label{fig:conf1-program}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l l}
  {\tt 0} &
  \multicolumn{2}{c}{
    \memory{3}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-14em}{0.75em}{a}
    \MemoryLabel{-10em}{0.75em}{b}
    \MemoryLabel{-7em}{0.75em}{c}
    \vspace{.5em}
  } \\
  {\tt 1} &
  \multicolumn{2}{c}{
    \memoryaddrs{12em}
    \memory{3}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-14em}{0.75em}{42}
    \MemoryLabel{-10em}{0.75em}{0}
    \MemoryLabel{-7em}{0.75em}{c}
    \vspace{.5em}
  } \\
  {\tt 10} &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf d}
 &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{a'}
  \MemoryLabel{-14em}{0.75em}{b'}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf d'}
  \\
  {\tt 11} &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf 0}
  &
  \memoryaddrs{20em}
  \memory{2}{\mainsealc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{a'}
  \MemoryLabel{-14em}{0.75em}{b'}
  \MemoryLabel{-10em}{0.75em}{2}
  \MemoryLabel{-7em}{0.75em}{\bf b'}
  \\
\end{tabular}

@@

\label{fig:conf1-trace}
\end{subfigure}
\caption{A Confidentiality Violation}
\end{figure}

Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
of the context of its call (except, of course, for its arguments).
We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
security. Noninterference describes knowledge: it states that for some component(s)\apt{??}
that we do not know, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the components by observing those results, violating
confidentiality.

In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt y}, which was not passed to it. The trace
in figure \ref{fig:conf1-trace} shows how noninterference works. \apt{Is this right? what is d?}
First, let \(a\),
\(b\), and \(c\) be the unknown initial values at their locations. On entry to {\tt f}
we again have 42 and 0 in {\tt x} and {\tt y}, and {\tt z} contains \(c\). We consider a
{\em variant state} that replaces the values of components that should be secret with
arbitrary values \(a'\), \(b'\), and \(c'\). After a step, we've copied 0 and \(c'\),
respectively, into {\tt z}. Since {\tt z} changed value, for confidentiality to hold,
its value should be the same in all variant states. But when \(c' \not = 0\), the
states change differently, which violates confidentiality. It would also be a violation
to print the secret value.\apt{all this is not too clear.}

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any components
whose values update and any printed values will be identical between the primary and variant
traces.

We alternatively could imagine a stepwise form of confidentiality in which we explicitly
track which addresses may be read (i.e., which have been initialized), and at each step
guarantee that the step alters state identically to a step from a variant state. Then we
wouldn't need to divide up the execution by calls. But this is further from the normal
framing of noninterference, requires explicitly tracking initializations, and would behave
no differently relative to our policies.\apt{not very clear.}

@@

>>>>

\section{Stack Safety, Intuitively}
\label{sec:running-example}
\colorlet{lgray}{gray!40}
\colorlet{lred}{red!40}
\colorlet{lblue}{blue!20}

\newcommand{\mainsealc}{cyan}
\newcommand{\fsealc}{green}
\newcommand{\unsealc}{lgray}
\newcommand{\emptyoutc}{white} % or gray for consistency with unusedc ?
\newcommand{\fulloutc}{white}
\newcommand{\badc}{lred}
\newcommand{\goodc}{lblue}

@@

Example \ref{fig:int1} shows an integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}, performs arithmetic, and returns a result.
In the initial memory layout the only frame is currently {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

@@

\begin{subfigure}{.33\textwidth}
\begin{verbatim}
int main() {
0:     int x,y;
1:     x = 42;
2:     y = 0;
3[C]:  y = f();
4[R]:  return x+y;
}

int f() {
10:     int w,z;
11:     w = 5;
12:     z = $(1,0);
13[R]:  return w+z;
}
\end{verbatim}
\label{fig:conf1-program}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memory{2}{\unsealc}
    \MemoryLabel{-6em}{0.75em}{42}
    \MemoryLabel{-3em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memory{2}{\unsealc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{0}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-7em}{0.75em}{42}
  \MemoryLabel{-3em}{0.75em}{0}
  &
  \memory{2}{\unsealc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{\(v_0\)}
  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
  \\
  {\tt 12} &
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-7em}{0.75em}{42}
  \MemoryLabel{-3em}{0.75em}{0}
  &
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{\(v_1\)}
  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
  \\
    {\tt 13} &
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{42}
  \MemoryLabel{-7em}{0.75em}{42}
  \MemoryLabel{-3em}{0.75em}{0}
  &
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  \hspace{1em}
  \memory{2}{\unsealc}
  \MemoryLabel{-17em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{\(v_2\)}
  \MemoryLabel{-7em}{0.75em}{\(v_2\)}
  \MemoryLabel{-3em}{0.75em}{\(v_3\)}
\end{tabular}

@@

\label{fig:conf1}
\end{subfigure}
\caption{A Confidentiality Violation}
\end{figure}

Meanwhile confidentiality intuitively states that the callee, {\tt f}, is unaware
of the context of its call (except, of course, for its arguments).
We state this in a style similar to {\em noninterference}, a well-known concept in theoretical
security. Noninterference describes knowledge: it states that for some system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the components by observing those results, violating
confidentiality.

In example \ref{fig:conf1} we see that the callee, {\tt f}, violates confidentiality
by reading {\tt main}'s variable {\tt x}, which was not passed to it. The trace
shows how noninterference applies to this scenario. At line 11, we create a {\em variant}
state with arbitrary values \(v_0\), etc., in place of the values that should
be hidden from {\tt f}. At line 12, we note that the callee's first variable, {\tt w}, has
changed to 5 in both the primary trace and the variant. The primary and variant agree on
the location that changed, so confidentiality was not violated. But after line 13,
the other variable, {\tt z}, has changed. In the primary trace it changed to 42, but
in the variant trace it might have any value \(v_2\). Say \(v_2 = 43\) -- then the variant
has behaved differently than the primary, violating confidentiality.

This form of confidentiality we will term {\em lockstep confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. Any system components
whose values update will be identical between the primary and variant traces. This includes not
just memory, but registers as well, so that returning a secret is a violation. And when we extend
the model with a notion of external output, it will include any values that are outputted.

We alternatively could imagine a stepwise form of confidentiality in which we explicitly
track which addresses may be read (i.e., which have been initialized), and at each step
guarantee that the step alters state identically to a step from a variant state. Then we
wouldn't need to divide up the execution by calls. But we can see in this example that taking
a single variant at the start of the call effectively captures initialization: when {\tt w} is
initialized, the primary and the variant agree on its value, so it can now be safely read.
This treatment is roughly in line with more standard uses of noninterference, which typically
vary states at the starts of whole program runs.

Together, stepwise integrity and lockstep confidentiality form the eager version of
local state encapsulation.

@@

It may be unclear in figure \ref{fig:conf1} why we also vary the initial values of {\tt w}
and {\tt z}. We will demonstrate by amending our toy model to make it closer to a real stack.
In figure \ref{fig:conf2} the topmost stack frame is no longer allocated as needed, but is an
infinite array of addresses. Variables are allocated and deallocated by moving the stack pointer,
and on a call, addresses to the left of the stack pointer are split off and pushed onto the
stack as sealed frames. On a return, the topmost sealed frame is prepended to the unsealed
frame. Now we can see how secrets might be left behind even in the unsealed portion of the stack.

\sna{Updated up to here.}
